---
title: "Fetal Mouse Hematology Data AIN93G- 35ppm Iron vs AIN-76A-100ppm Iron"
author: "Kaylee Helfrich"
date: "2/20/2020"
output: 
  pdf_document: default
  html_document: default
data: "Fetal_Mouse_Hematology_CLEAN_5Groups_93G&76A.csv"
---

Built with R `r getRversion()`

```{r RMarkdown setup, include = FALSE, message = FALSE}
require(knitr)
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60), tidy=TRUE)
knitr::opts_knit$set(root.dir = "~/KayleeStuff/Smith_Lab/Data/Mouse_Validation_Data/Hematology")
options(tinytex.verbose = TRUE)
```

# Graphing of CBCs
This document details my graphing of the hematology data from fetal CBCs. The AIN-93G-35ppm iron groups were collected in the fall of 2019 by Brandon Presswood, Dane Ricketts, Olivia Rivera, and Katie Walter. The design was the standard design that we use, with 3.0g/kg ethanol from GD8.5-17.5 or 80% isocaloric maltodextrin. All dams consumed AIN-93G diet from 3 weeks prior to pregnancy through pregnancy and collection. Tissues were harvested on GD17.5, 4 hours after the final dose. Blood was collected from the fetuses by decapitation and heparin collection on the inverted fetal trunk. Blood was pooled across the litter. The blood was run using the diluted mode on the pochi analyzer. In contrast, the AIN-76A- 100ppm iron groups were collected in the spring of 2019 (March-April) by Kaylee Helfrich. The design was standard, except that dams consumed AIN-76A diet containing 100ppm of iron instead of 35ppm. In addition, one of the alcohol groups received its last dose of alcohol on GD16.5 and was collected 24hours following this final dose. All other aspects of this study remained the same as our standard model. All graphs are mean with standard deviation, since there are not multiple measures per dot. This analysis will allow us to compare the 5 groups, and potentially narrow down the analysis later to just 2 or 4 groups, depending upon the question.

```{r include = FALSE}
rm(list=ls())
setwd("~/KayleeStuff/Smith_Lab/Data/Mouse_Validation_Data/Hematology")
library(ggplot2)
library(plotrix)
```

```{r Import file and rearrange by treatment to be alphabetical}
File <- "Fetal_Mouse_Hematology_CLEAN_5Groups_93G&76A.csv"
data <- read.csv(File, header=TRUE) 
data_frame <- data.frame(data)
data_frame <- data_frame[order(data_frame$Treatment), ]
```
```{r echo = FALSE}
print(data_frame)
```
```{r Prepare data for graphing}
ActualGroupOrder <- c("MD-100ppm", "MD-35ppm", "PAE-100ppm", "PAE-100ppm-24h", "PAE-35ppm")
NumPoints <- 9 #Make sure to set the integer to the number of dots/values you have for each treatment
```

# White Blood Cells
## All code is included for this first example, but thereafter, only the heading and graphs are shown. For all graphs, all data is graphed unless mentioned otherwise. Messages and warnings are excluded, since NA's were included to make all groups even, and I didn't want the code to print out these errors when they weren't necessary.
```{r, warning = FALSE, message = FALSE}
#Mean- calculate mean for column, repeat it to number of points, and create dataframe
Mean <- tapply(data_frame$WBC, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change selected column to change data for graph
Mean <- as.data.frame(rep(Mean, each = NumPoints)) #repeat mean a set number of times
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) #repeat treatment a set number of times
colnames(Mean)[1] <- "Mean" #fix column name

#Stdev- calculate standard deviation for column, repeat it to number of points, and create dataframe
Stdev <- tapply(data_frame$WBC, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change selected column
Stdev <- as.data.frame(rep(Stdev, each = NumPoints)) #repeat stdev a set number of times
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints) #repeat treatment a set number of times
colnames(Stdev)[1] <- "Stdev" #fix column name

#Combine mean, standard deviation, and individual data points
dataframe_combine <- cbind(Mean, Stdev, data_frame$WBC) #make sure to change selected column
colnames(dataframe_combine)[5] <- "Value" #Fix name of individual data point column
dataframe_combine <- dataframe_combine[-2] #Remove extra treatment column

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev #upper error
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev #lower error

#Make Graph#
png("Fetal_WBCs_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC",
                             "PAE-35ppm" = "#6699FF",
                             "MD-100ppm" = "#660066",
                             "PAE-100ppm" = "#990099",
                             "PAE-100ppm-24h" = "#CC33FF"
                             )) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,12)) +
  ggtitle("Fetal WBCs") +
  ylab(expression("Fetal WBCs (x10^3 cells/uL)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r White Blood Cells, echo = FALSE}
print(p)
```

# Red Blood Cells
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$RBC, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$RBC, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$RBC) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_RBCs_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,4)) +
  ggtitle("Fetal RBCs") +
  ylab(expression("Fetal RBCs (x10^6 cells/uL)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Red Blood Cells, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Hemoglobin
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$HGB, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$HGB, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$HGB) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_Hemoglobin_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,15)) +
  ggtitle("Fetal Hemoglobin") +
  ylab(expression("Fetal Hemoglobin (g/dL)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Hemoglobin, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Hematocrit
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$HCT, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$HCT, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$HCT) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_Hematocrit_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,50)) +
  ggtitle("Fetal Hematocrit") +
  ylab(expression("Fetal Hematocrit (%)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Hematocrit, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Mean Cell Volume
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$MCV, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$MCV, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$MCV) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_MeanCellVolume_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,150)) +
  ggtitle("Fetal Mean Cell Volume") +
  ylab(expression("Fetal MCV (fL)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Mean Cell Volume, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Mean Cell Hemoglobin
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$MCH, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$MCH, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$MCH) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_MeanCellHemoglobin_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,40)) +
  ggtitle("Fetal Mean Cell Hemoglobin") +
  ylab(expression("Fetal MCH (pg)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Mean Cell Hemoglobin, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Mean Cell Hemoglobin Concentration 
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$MCHC, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$MCHC, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$MCHC) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_MeanCellHemoglobinConcentration_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,30)) +
  ggtitle("Fetal Mean Cell Hemoglobin Concentration") +
  ylab(expression("Fetal MCHC (g/dL)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Mean Cell Hemoglobin Concentration, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Platelets
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$PLT, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$PLT, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$PLT) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_Platelets_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,160)) +
  ggtitle("Fetal Platelets") +
  ylab(expression("Fetal Platelets (x10^3 platelets/uL)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Platelets, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Zinc Protoporphyrin
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Mean
Mean <- tapply(data_frame$ZnPP, as.character(data_frame$Treatment), mean, na.rm=TRUE) #make sure to change 
Mean <- as.data.frame(rep(Mean, each = NumPoints))
Mean$Treatment <- rep(ActualGroupOrder, each = NumPoints) 
colnames(Mean)[1] <- "Mean"

#Stdev
Stdev <- tapply(data_frame$ZnPP, as.character(data_frame$Treatment), sd, na.rm=TRUE) #make sure to change 
Stdev <- as.data.frame(rep(Stdev, each = NumPoints))
Stdev$Treatment <- rep(ActualGroupOrder, each = NumPoints)
colnames(Stdev)[1] <- "Stdev"

#Combine
dataframe_combine <- cbind(Mean, Stdev, data_frame$ZnPP) #make sure to change
colnames(dataframe_combine)[5] <- "Value"
dataframe_combine <- dataframe_combine[-2]

#Error bars
dataframe_combine$Upper <- dataframe_combine$Mean + dataframe_combine$Stdev
dataframe_combine$Lower <- dataframe_combine$Mean - dataframe_combine$Stdev

#Make Graph#
png("Fetal_ZincProtoporphyrin_5Groups_Dotplot.png", units="in", width=7, height=7, res=600) 

sets <- c("MD-35ppm", "PAE-35ppm", "MD-100ppm", "PAE-100ppm", "PAE-100ppm-24h")

p<- ggplot(dataframe_combine, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", color="black", aes(fill = Treatment),
           position=position_dodge()) +
  scale_fill_manual(values=c("MD-35ppm" = "#0000CC", 
                             "PAE-35ppm" = "#6699FF", 
                             "MD-100ppm" = "#660066", 
                             "PAE-100ppm" = "#990099", 
                             "PAE-100ppm-24h" = "#CC33FF")) +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,
                position=position_dodge(.9)) + 
  geom_point(aes(x=Treatment, y=Value),
             size = 1.25, shape = 19, position = position_jitter(width=.15, height=0.1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,800)) +
  ggtitle("Fetal Zinc Protoporphyrin") +
  ylab(expression("Fetal ZnPP (umol ZnPP/mol heme)")) + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
        axis.title.x = element_blank(), #gets rid of x-axis label
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), #gets rid of major gridlines
        panel.grid.minor = element_blank(), #gets rid of minor gridlines
        panel.background = element_blank(), #turns background white instead of gray
        axis.line = element_line(colour = "black"), 
        legend.position="none", #gets rid of legend
        axis.text=element_text(size=18, color = "black"), #sets size of x and y axis labels
        axis.text.x=element_text(angle = 45, vjust = .95, hjust = 0.9),
        axis.title=element_text(size=20,face="bold")) + 
  scale_x_discrete (limits = sets)
print(p)
dev.off() 
```
```{r Zinc Protoporphyrin, echo = FALSE, warning = FALSE, message = FALSE}
print(p)
```

# Session Information
```{r, echo = FALSE}
session_info <- devtools::session_info()
print(session_info)
```